---
title: "DESeq2 Up/Down Summary and Plots"
author: "Chuck Roesel"
date: "10/28/2017"
output: html_document
---
##Instructions
Write an R script named upDownSummary.R and add the code shown below. Run line-by-line and check the output.

##Load Libraries
```{r Load Libraries, message=FALSE, warning=FALSE, paged.print=FALSE}
library("DESeq2")
library("RColorBrewer")
library("pheatmap")
```

##Load Summarized DE Results
```{r}
load("summarizedDeResults.RData")
```

##Plot Dispersion
Review information on Dispersion in the [DESeq2 manual](https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#dispersion-plot-and-fitting-alternatives)
```{r}
plotDispEsts(ddsAll)
```

##Get Summaries
Get subsets of genes up-regulated or down-regulated for each factor using a p-value cutoff of .05.
```{r, echo=TRUE}
logThreshold=0
alpha=.05
upVibrio <- subset(dfBothAnnot, (padj.Vibrio < alpha) & (lfc.Vibrio > logThreshold))
downVibrio <- subset(dfBothAnnot, (padj.Vibrio < alpha) & (lfc.Vibrio < logThreshold))
upMenthol <- subset(dfBothAnnot, (padj.Menthol < alpha) & (lfc.Menthol > logThreshold))
downMenthol <- subset(dfBothAnnot, (padj.Menthol < alpha) & (lfc.Menthol < logThreshold))
upInter <- subset(dfBothAnnot, (padj.Inter < alpha) & (lfc.Inter > logThreshold))
downInter <- subset(dfBothAnnot, (padj.Inter < alpha) & (lfc.Inter < logThreshold))
```

##Vibrio Volcano Plot
Plot log2 fold change as a function of mean normalized counts. Each point on the graph represents a gene. Red indicates an adjusted p-value of .05 or lower.
```{r, echo=TRUE}
DESeq2::plotMA(resVibrio, ylim=c(-10,10), alpha=.05)
```

##Menthol Volcano Plot
Plot log2 fold change as a function of mean normalized counts.
```{r, echo=TRUE}
DESeq2::plotMA(resMenthol, ylim=c(-10,10), alpha=.05)
```

```{r, echo=TRUE}
DESeq2::plotMA(resInter, ylim=c(-10,10), alpha=.05)
```


##Principal Component Analysis
Perform principal component analysis (PCA) to determine percentage of variation explained by each
varable. About 40% explained by the first component, and about 13% by the second component.

```{r}
plotPCA(rldAll, intgroup=c("Vibrio", "Menthol"))
```

##Venn Diagram
Make a Venn diagram showing overlap between Vibrio Up, Vibrio Down, Menthol Up, Menthol Down.
```{r}
upVibrio$count <- rep(1,nrow(upVibrio))
downVibrio$count <- rep(1, nrow(downVibrio))
upMenthol$count <- rep(1, nrow(upMenthol))
downMenthol$count <- rep(1, nrow(downMenthol))
upInter$count <- rep(1, nrow(upInter))
downInter$count <- rep(1, nrow(downInter))
vibrioCount <- merge(upVibrio, downVibrio, by=0, all=TRUE, suffixes=c("Up", "Down"))
mentholCount <- merge(upMenthol, downMenthol, by=0, all=TRUE, suffixes=c("Up", "Down"))
interCount <- merge(upInter, downInter, by=0, all=TRUE, suffixes=c("Up", "Down"))
rownames(vibrioCount) <- vibrioCount$Row.names
rownames(mentholCount) <- mentholCount$Row.names
rownames(interCount) <- interCount$Row.names
vibrioCount <- subset(vibrioCount, select = -c(Row.names) )
mentholCount <- subset(mentholCount, select = -c(Row.names) )
interCount <- subset(interCount, select = -c(Row.names) )
bothCount <- merge(vibrioCount, mentholCount, by=0, all=TRUE, suffixes=c("Vibrio", "Menthol"))
rownames(bothCount) <- bothCount$Row.names
bothCount <- subset(bothCount, select = -c(Row.names) )
bothCount <- merge(bothCount, interCount, by=0, all=TRUE)
rownames(bothCount) <- bothCount$Row.names
bothCount <- subset(bothCount, select = -c(Row.names) )
bothCount <- subset(bothCount, select = c(countUpVibrio, countDownVibrio, countUpMenthol, 
                                          countDownMenthol, countUp, countDown) )
head(bothCount)
```

###Change NAs to zero
```{r}
bothCount$countUpVibrio[is.na(bothCount$countUpVibrio)] <- 0
bothCount$countDownVibrio[is.na(bothCount$countDownVibrio)] <- 0
bothCount$countUpMenthol[is.na(bothCount$countUpMenthol)] <- 0
bothCount$countDownMenthol[is.na(bothCount$countDownMenthol)] <- 0
bothCount$countUp[is.na(bothCount$countUp)] <- 0
bothCount$countDown[is.na(bothCount$countDown)] <- 0
```

##Display Venn Diagram
```{r}
library("limma")
vennDiagram(vennCounts(subset(bothCount, select=c(countUpVibrio, countDownVibrio, countUpMenthol, countDownMenthol, countUp))), names=c("UpVibrio", "DownVibrio","UpMenthol", "DownMenthol", "UpInter"), circle.col=c("red", "green", "blue", "yellow", "coral2"))
```
```{r}
#install.packages("UpSetR")
library(UpSetR)
listInput = list(UpVibrio = rownames(upVibrio), UpMenthol= rownames(upMenthol), DownVibrio=rownames(downVibrio), DownMenthol=rownames(downMenthol), UpInteraction=rownames(upInter), DownInteraction=rownames(downInter))
#listInput
```

```{r}
upset(fromList(listInput), order.by = "freq", nsets=6)
```
##RLog Transformation
Perform an rlog transformation on the counts to construct a sample distance matrix. The sample distance matrix provides an measure of how similar overall gene expression patterns are for each sample.
```{r}
rldAll <- rlog(ddsAll, blind=FALSE)
sampleDists <- dist(t(assay(rldAll)))
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(rldAll$Vibrio, rldAll$Menthol, sep="-")
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
```

##Cluster samples based on overall gene expression patterns
```{r}
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
```
